Expected Dependency Pair Match:\\ Predicting {HTER} improvements with expected syntactic structure

### 2009-05-10 mtjournal.tex
new macros for tree diagrams
beginning of the paradigm section (with implementation details)

\usepackage{graphicx}

\newcommand{\arclabel}[1]{\ensuremath{\stackrel{#1}{\to}}}
\newcommand{\precision}[1]{\ensuremath{\textrm{Prec}\left(#1\right)}}
\newcommand{\recall}[1]{\ensuremath{\textrm{Recall}\left(#1\right)}}

@@

We may not completely trust the parser's best parse.  The parser
itself, if we use a probabilistic parser, can provide an $n$-best
parse list forest for the translation reference and translation
hypothesis. We use the probability statistics of the forest to compute
expected counts for each decomposition. Though this approach yields
partial counts, standard comparisons like precision and recall are
still valid.

@@

@@

\begin{itemize}
\item Listing possible component valences
  \begin{itemize}
  \item dependent + labeled link + head 
  \item outbound only (dependent +
    label)
  \item inbound only (label + head)
  \item word-word (skipping label)
  \end{itemize}
  \begin{itemize}
  \item 1-grams
  \item 2-grams
  \item \ldots{}
  \end{itemize}

>>>>

\usepackage{graphicx}
\usepackage{qtree}

%\newcommand{\headlabel}[2]{\ensuremath{\underset{\textrm{#2}}{\textrm{#1}}}}
\newcommand{\headlabel}[2]{
  \begin{tabular}{c} #1\\
    {\small #2} \end{tabular}}
\newcommand{\arclabel}[1]{\ensuremath{\stackrel{#1}{\to}}}

\newcommand{\precision}[1]{\ensuremath{\textrm{Prec}\left(#1\right)}}
\newcommand{\recall}[1]{\ensuremath{\textrm{Recall}\left(#1\right)}}

@@

We may not completely trust the parser's best parse.  The parser
itself, if we use a probabilistic parser, can provide an $n$-best
parse list forest for the translation reference and translation
hypothesis. We use the probability statistics of the forest to compute
expected counts for each decomposition. Though this approach yields
partial counts, standard comparisons like precision and recall are
still valid, and using the expectations can help cope with both error
in the parser and ambiguity in the translations (reference and
hypothesis).

@@

In principle, the family of DPM measures may be implemented with
any parser that generates a dependency graph (a single labelled arc
for each word, pointing to its head-word). Prior work
\cite{owczarzak07labelleddepseval} on related
measures has used an LFG parser \cite{cahill04lfg} or
an unlabelled dependency tree \cite{liu05syntaxformteval}.

%
In this work, we use a state-of-the-art PCFG (the first stage of
\cite{charniak-johnson:2005:ACL}) and context-free head-finding rules
\cite{magerman95headfinding} to generate a 50-best list of dependency
trees for each hypothesis and reference translation.  We use the
parser's default (English-language) Wall Street Journal training
parameters.  Head-finding uses the Charniak parser's rules, with three
modifications: prepositional and complementizer phrases choose nominal
and verbal heads respectively (rather than functional heads) and
auxiliary verbs are modifiers of main verbs (rather than the
converse).

Having constructed the dependency tree, we label the arcs as $d
\stackrel{A/B}{\to} h$, where the arc label $A/B$ between dependent
$d$ and its head $h$ is composed of $A$ (the lowest constituent-label
headed by $h$ and dominating $d$) and $B$ (the highest constituent
label headed by $d$). 
\begin{figure}
  % \centering
  \Tree
  [.\headlabel{S}{stumbled}
    [.\headlabel{NP}{cat}
      [.\headlabel{DT}{the} the ]
      [.\headlabel{NN}{cat} cat ]
    ]
    [.\headlabel{VP}{stumbled} 
      [.\headlabel{VBD}{stumbled} stumbled ]
    ]
  ]
  \\
  \begin{center}
    \includegraphics[scale=0.6]{dpm-example-depextract.eps}
  \end{center}
  \caption{An example constituent tree (heads of each constituent are
    listed small below the label) and the labelled dependency tree
    derived from it using the strategy described in
    section~\ref{sec:paradigm}.}
  \label{fig:depextract}
\end{figure}
For example, in figure~\ref{fig:depextract}, the S node is the
lowest node headed by \emph{stumbled} that dominates \emph{cat}, and
the NP node is the highest constituent label headed by \emph{cat}, so
the arc between \emph{cat} and \emph{stumbled} is labelled \arclabel{S/NP}.
%
This strategy is very similar to one adopted in the reference
implementation of labelled-dependency \textsc{SParseval}, and may be
considered as an approximation of the rich semantics generated by
\cite{cahill04lfg} or another heavily knowledge-engineered
parser, but with much less knowledge-engineering required.

The $A/B$ labels are not as descriptive as the LFG semantics, but they
have a similar resolution, e.g.\ the $\stackrel{S/NP}{\to}$ arc label
usually represents a subject dependent of a sentential verb.

In the experiments presented later in this article, we explore the
space of metrics outlined above along multiple dimensions:
\begin{description}
\item[Decompositions]: we consider syntactic decompositions of the
  following types.
  \begin{description}
  \item[dlh] $\left\langle \textrm{Dependent}, \textrm{arc Label},
      \textrm{Head}\right\rangle$ here, every tuple includes two words
    and the dependency-label between them.
  \item[dl] $\left\langle \textrm{Dependent}, \textrm{arc Label}
      \right\rangle$ the tuple is the word plus the label of the arc
      linking it to its head; this decomposition marks how the word
      fits into its syntactic context (what it modifies).
    \item[lh] $\left\langle \textrm{arc Label}, \textrm{Head}
      \right\rangle$ each tuple here is the dependency relation plus
      the word being modified; this decomposition implicitly marks how
      key the word is to the sentence.
    \item[dh] $\left\langle \textrm{Dependent}, \textrm{Head}
      \right\rangle$ here, drops the arc-label. This decomposition
      ignores the syntactic-role information implied in the arc labels.
    \item[1g,2g] these decompositions are syntactically-null
      they just includes each 1-gram (or 2-gram) without any
      syntactic labeling at all.  They are thus simple measures of
      unigram (bigram)
      precision and recall.
  \end{description}
\item[Forest size]
Although the parser can generate considerably more, we consider forests of size 1 and of size 50
\item[Forest confidence]
\end{description}

@@

\begin{itemize}
